---
title: "Chicago Crime"
author: "Xin Guan, Vera Hudak, Yuqi Zhang"
date: "2023-11-24"
output:
  pdf_document: default
  html_document: default
---
```{r}
# Packages required
library(lubridate)
library(dplyr)
library(magrittr)
library(tidyverse)
library(ggplot2)
library(caret)
library(pROC)
library(MLmetrics)
library(ROSE)
```

## Data Processing

```{r}
# Read data
Crime_data <- read.csv("OriginalData.csv")
```

### Processing variable 'Date'
```{r}
# We want to firstly convert the format of date for the `lubridate` package, then extract information from the date to create new variables.
Crime_data$newDate <- mdy_hms(Crime_data$Date)
Crime_data$Hour <- hour(Crime_data$newDate)
Crime_data$WeekDay <- weekdays(Crime_data$newDate)
Crime_data$DayOfMonth <- day(Crime_data$newDate)
Crime_data$DayOfYear <- yday(Crime_data$newDate)
Crime_data$Month <- month(Crime_data$newDate, label = TRUE, abbr = FALSE)
Crime_data$Time <- hour(Crime_data$newDate)*100 + minute(Crime_data$newDate) #This format of `Time` variable contains information of hour and minute only
Crime_data$TimeOfDay <- cut(
  hour(Crime_data$newDate),
  breaks= c(-Inf, 5, 12, 17, 20, Inf),
  labels = c("Night", "Early Morning", "Morning", "Afternoon", "Evening"),
  include.lowest = TRUE
)
```

### Processing missing data
```{r}
colSums(is.na(Crime_data))
```
We can see that the number of missing values for the variables `X.Coordinate`, `Y.Coordinate`, `Latitude`, and `Longtitude` are exactly the same, we can deduce that the coordinates are calculated from the latitude and longitude, so we don't need to include both pairs of location information. Also since the number of missing value is small compare to the number of data size, we will just eliminate the rows with missing values.
```{r}
Crime_data %<>%  na.omit()
```

## Expalnatory Data Analysis








## Feature Selection
Our first task is to predict whether arrest or not given time, location, and the crime type.
```{r}
Binary_pred_df <- Crime_data %>% select(c("ID","X.Coordinate", "Y.Coordinate","Hour","Time","WeekDay","DayOfMonth","DayOfYear","Primary.Type","Arrest"))
#???location variables
```

## Imbalance Data Experiments








## Baseline Model

```{r}
#make this example reproducible
set.seed(1)

#use 80% of dataset as training set and 30% as test set 
train <- Binary_pred_df %>% dplyr::sample_frac(0.80)
train$Arrest <- as.factor(train$Arrest)
test  <- dplyr::anti_join(Binary_pred_df, train, by = 'ID')
test$Arrest <- as.factor(test$Arrest)
```

```{r}
# Define a 5-fold cross validation
ctrl <- trainControl(method = "cv", number = 5, summaryFunction = twoClassSummary, classProbs = TRUE)
```

```{r}
base_m <- train(Arrest ~ X.Coordinate + Y.Coordinate + DayOfYear + Hour + 
    Primary.Type, data = train, method = "glm", family = "binomial", trControl = ctrl, metric = "ROC")
```


```{r}
pred <- predict(base_m, test[,c("X.Coordinate", "Y.Coordinate", "DayOfYear", "Hour", "Primary.Type")], type = "prob")
```

```{r}
binary_predictions <- ifelse(pred[,2] >= 0.5, "true", "false")
# create confusion matrix 
confusionMatrix(as.factor(binary_predictions), test$Arrest,
                mode = "everything",
                positive="true")
```

```{r}
# Create a ROC curve object
roc_curve <- roc(test$Arrest, pred[,2])

# Plot the ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)

# Add AUC to the plot
auc_value <- auc(roc_curve)
text(0.8, 0.2, paste("AUC =", round(auc_value, 3)), col = "blue", cex = 1.2)
```
